---
title: "Crashes in California (2014-2023)"
format:
  closeread-html:
    cr-style:
      section-background-color: white
      narrative-background-color-sidebar: black
      narrative-text-color-sidebar: white
      narrative-font-family: 'Georgia, "Times New Roman", Times, serif'
      narrative-font-size: 1.8rem
embed-resources: true
---

```{r setup, include=FALSE}
library(tidyverse)

crashes = read_csv("crashes_per_year.csv")
```


Crashes per year in California

::::{.cr-section}

Have you ever wondered how many crashes per year occur in California? @cr-plot1

:::{#cr-plot1}
```{r plot1}
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = COUNT)) +
  geom_col(fill = "#FFFFFF00") +
  scale_x_continuous(breaks = 2014:2023) +
  labs(x = "") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```
:::


Here's the distribution of crashes in California between 2014 and 2023 @cr-plot2

:::{#cr-plot2}
```{r plot2}
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = COUNT)) +
  geom_col(fill = "#AAAAAA") +
  scale_x_continuous(breaks = 2014:2023) +
  labs(x = "") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```
:::

The number of crashes has always exceeded 150,000 ... @cr-plot3

:::{#cr-plot3}
```{r plot3}
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = COUNT)) +
  geom_col(fill = "#AAAAAA") +
  geom_hline(yintercept = 150000) +
  scale_x_continuous(breaks = 2014:2023) +
  labs(x = "") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```
:::


... except in 2020 when the Covid pandemic hit @cr-plot4

:::{#cr-plot4}
```{r plot4}
crashes |> 
  ggplot(aes(x = ACCIDENT_YEAR, y = COUNT)) +
  geom_col(fill = c(rep("#AAAAAA", 6), "tomato", rep("#AAAAAA", 3))) +
  geom_hline(yintercept = 150000) +
  scale_x_continuous(breaks = 2014:2023) +
  labs(x = "") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```
:::

Why did the number of crashes in 2021, 2022, and 2023 remain below the pre-pandemic
levels? [@cr-plot4]{pan-to="-25%,25%" scale-by="1.5"}

Probably because of: high gas prices, and working from home @cr-plot4

::::
